\clearpage
\begin{centering}
\textbf{ACKNOWLEDGEMENTS}\\
\vspace{\baselineskip}
\end{centering}

%Insert your dedication text here
First, I would like to thank my advisor, Professor Alberto Fern\'andez-Nieves; it has been an incredible experience working with him over these past six years.
Alberto is a truly fantastic mentor and has been a constant source of guidance and inspiration.
He continuously pushes me and the rest of the group to be the best versions of ourselves possible, and his influence is the main reason I have become the scientist that I am.
I would also like to thank the rest of faculty members who have agreed to serve on my committee: Professor Paul Goldbart, Professor Peter Yunker, Professor Brian Kennedy, and Professor Eric Weeks.
For the wonderful opportunity of being paid to learn and be a student, thank you to the Georgia Tech School of Physics and to the FLAMEL program.

This work was truly the product of collaboration, and I would especially like to thank those who directly contributed to this dissertation.
Professor Luca Giomi and Daniel Pearce performed the numerical simulations on the active nematic toroids presented in Chapter 3.
Karthick Nayani started the work on homeotropic nematic toroids and performed some of the experiments presented in Chapter 4.
Jayalakshmi Vallamkondu started the work on the nematic bridges and performed some of the experiments presented in Chapter 5.
Susannah Klane\u{c}ek, an undergraduate in our group, performed many of the experiments on the nematic bridges; Chapter 5 would not exist without her.
Professor Paul Goldbart and Shengnan Huang performed the numerical calculations for the nematic bridges.

To the Fern\'andez-Nieves Lab, past and present, thank you for your inspiration, support, conversation, and so much more.
Alexandros Fragkopoulos for helping me with homework in our first year, for serving as a sounding board for many ideas and thoughts over the years, and for being a wonderful friend.
Karthik Nayani for many discussions about liquid crystals and unrelated tangents.
Ya-wen (Winnie) Chang for her help and wisdom both in and out of the lab.
Mike Dimitriyev for his patience in teaching me math and so much more.
The rest of my labmates, Ekapop, John, Miguel, Josefa, Mike T., Caleb, Jonas, Sam, and Boyang, thank you for more than you know.
It has been a pleasure working with all of you. \\

\noindent Finally, to my family.
Mama, Baba, Reilly, Clayton, and Kristine, thank you for your unending love and support.
This one's for you.

\clearpage
%\pagenumbering{gobble}  %remove page number on summary page
